Uncertainty Estimation Methods for Countering Attacks on Machine-Generated Text Detectors

### 2025-03-29 paper/main.tex
small changes

@@

\begin{figure}[bhtp]
	\includegraphics[width=\textwidth]{plot1.png}
	\caption{Perplexity distribution differs for different models.}
	\label{fig:schema}
\end{figure}

We processed texts from both labels using the LLM (Llama-3.1-8B-Instruct) and utilized contextual logits to compute two metrics: perplexity and entropy. By analyzing the distribution of points on the plane (Figure 1), a linear relationship between perplexity and entropy becomes evident. It is noticeable that the distribution of models is very different. Davinci and dolly have similar entropy as humans, while chatGPT and gpt4 is very different. This finding is promising, as it suggests that the distinction between MG and HW text distributions can potentially be detected using uncertainty estimation (UE). In the following sections, we explore this idea further to develop a more robust detection framework.

>>>>

As such metrics, we will use \textbf{perplexity}: 
$$PPL  = \exp\left(-\frac{1}{L} \sum_{l=1}^{L} \log P(w_l | w_{<l})\right)$$

And also \textbf{mean token entropy}, where we simply average entropy of each individual token in the generated sequence.

$$H = -\frac{1}{N} \sum_{i=1}^{N} \sum_{j} P(w_j | w_{<i}) \log P(w_j | w_{<i})$$

@@

\begin{figure}[bhtp]
	\includegraphics[width=\textwidth]{arxiv.png}
	\caption{Perplexity distribution differs for different models.}
	\label{fig:schema}
\end{figure}

We processed texts from both labels using the LLM (Llama-3.1-8B-Instruct) and utilized contextual logits to compute two metrics: perplexity and entropy. For consistency of the experiment, all data was taken from ARXIV. By analyzing the distribution of points on the plane (Figure 1). It is noticeable that the distribution of models is very different. The texts of different models are clustered according to the values of perplexity, which is a very interesting result. Dolly has similar entropy as humans, while chatGPT, bloomz, davinci and gpt4 is very different. This finding is promising, as it suggests that the distinction between MG and HW text distributions can potentially be detected using uncertainty estimation (UE). In the following sections, we explore this idea further to develop a more robust detection framework.
